Step by Step: Using Step Selection Functions to Simulate Dispersal and Assess Landscape Connectivity

### 2021-07-01 Chapter_1/04_Manuscript/Manuscript.tex
Updated graphics and discussion

% Previous Paper
\subsection{Previous Paper (90\%)}
In a previous study, we assessed landscape connectivity for dispersing African
wild dogs within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}.
Specifically, we fitted a basic habitat selection model and predicted a
permeability surface that we used to compute least-cost paths and corridors. We
now expand on this knowledge and use ISSFs to develop a more mechanistic
movement model of dispersing wild dogs (\Cref{GraphicalAbstract}). We employ the
model to simulate dispersers moving across the KAZA-TFCA and generate three
distinct connectivity maps, each shedding light onto a different aspect of
connectivity. With this work, we exemplify how ISSFs can be utilized for
dispersal simulations and we discuss several benefits of this approach over
traditional connectivity modeling techniques such as least-cost analyses and
circuit theory. Most importantly, simulations based on ISSFs provide a more
generic view on how connectivity emerges and to which degree connectivity
depends on the dispersal duration. In addition, by generating proper dispersal
trajectories, network theory can be applied to calculate network metrics that
are pertinent to connectivity analysis.

@@

Besides habitat covariates, we computed movement metrics that we used as
movement covariates in our models. Movement metrics were calculated for each
step and included the step length (\textsf{sl}), its natural logarithm
(\textsf{log(sl)}), and the cosine of the relative turning angle
(\textsf{cos(ta)}) (for details see \citep{Avgar.2016, Fieberg.2020}). Because
wild dogs follow a diurnal activity pattern, we also coded a binary variable
(\textsf{LowActivity}) indicating whether a step was realized during periods of
low wild dog activity (17:00 to 09:00 local time) or high wild dog activity
(09:00 to 17:00 local time). Handling and manipulation of all data, as well as
all models and simulations were implemented with the statistical software {\tt
R}, version 3.6.6 \citep{R.2019}. Several helper functions were written in {\tt
C++} and imported into {\tt R} using the {\tt Rcpp} package
\citep{Eddelbuettel.2011, Eddelbuettel.2013}

@@

\subsection{Convergence (50\%)}
Our examination of the traversal frequency as a function of the number of
simulated dispersers shows that already after a few simulations the mean
traversal frequency remains stable and on average changes only little when
adding further dispersers (\Cref{Convergence} (a) and (b)). While variability
keeps decreasing with additional dispersers, the marginal benefit of having
further dispersers decreases following a negative-exponential trend
(\Cref{Convergence} (c)).

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{99_Convergence}
    \caption{Relative traversal frequency across 1'000 checkpoints (5 km x 5)
    distributed in the study area. The relative traversal frequency is plotted
    against the number of simulated individuals to visualize how quickly the
    metric converges to a steady state. (a) Replicated (100 times) relative
    traversal frequencies across three randomly chosen checkpoints as well as
    the corresponding 95\% prediction interval (PI). (b) Averaged relative
    traversal frequency across all checkpoints and replicates including a 95\%
    PI. (c) Width of the PI in relation to the number of simulated dispersers.}
    \label{Convergence}
  \end{center}
\end{figure}

@@

\section{Discussion}
\subsection{Short Summary (90\%)}
% Maybe put this at the end?
We used ISSFs to analyse data of dispersing wild dogs and parametrize a fully
mechanistic movement model describing how dispersers move through the available
landscape. We employed the parametrized model as an individual-based movement
model and simulated 80'000 dispersing wild dogs moving 2'000 steps across the
extent of the KAZA-TFCA, the world's largest transboundary conservation area.
Based on simulated dispersal trajectories, we derived a set of three
complementary maps, all geared towards a better understanding of dispersal and
landscape connectivity. The set of maps included a heatmap, revealing frequently
traversed areas, a betweenness-map, delineating critical dispersal corridors and
bottlenecks, and a map of inter-patch connectivity, indicating the presence or
absence of functional links between national parks. We thereby showcase how
ISSFs can be utilized as simple, yet powerful framework to parametrize movement
models and simulate animal movement with the aim of assessing landscape
connectivity. Such an individual-based simulation overcomes several conceptual
shortcomings inherent to more traditional connectivity modeling techniques, such
as least-cost analyses and circuit theory. Nevertheless, we suggest to view
similar individual-based simulations as complement, and not as substitute, to
traditional connectivity models.

\subsection{Movement Model (80 \%)}
Our movement model of dispersing wild dogs comprised a habitat kernel, a
movement kernel, and their interactions. Together, the kernels described habitat
and movement preferences of dispersing individuals, as well as how movement
behavior was affected by habitat conditions. Parameter estimates from the
habitat kernel revealed that dispersing wild dogs avoid water, prefer its
proximity, avoid woodland, prefer shrubs/grassland, and avoid areas dominated by
humans. These findings are consistent with the findings of previous studies on
dispersing wild dogs \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019,
Oneill.2020}, as well as with an earlier dispersal model that we developed
to investigate the habitat kernel of dispersers \citep{Hofmann.2021}.

By expanding our initial model to include a proper movement kernel, we were able
to model several additional complexities inherent to dispersal. It is, for
instance, well known that dispersers typically move with directional persistence
and that step lengths are correlated with turning angles. That is, larger steps
usually coincide with smaller turning angles and vice versa (cite someone).
While this correlation structure could be rendered by sampling turning angles
and step lengths from copula probability distributions \citep{Hodel.2021}, the
ISSF framework allowed us to model similar behavior directly in the movement
model. Besides accounting for directional persistence, we further exploited the
ISSF framework and included interactions between movement covariates to
represent the fact that wild dogs mainly move during the more temperate morning
and evening hours but tend to rest during the remainder of the day. The final
model thus rendered a pulse-like movement behavior, where dispersers realize
larger steps during morning and evening hours, but shorter steps during the rest
of the day.

By including interactions between habitat and movement covariates, we also
acknowledged that movement behavior of dispersers depends on habitat conditions.
For example, the final model contained an interaction between water cover and
step length, highlighting that dispersers are more likely to realize short steps
(i.e. move slower) in areas covered by water. Likewise, the interaction between
water cover and turning angles revealed that dispersers move less directional
when moving across water bodies. We believe that this behavior is owed to the
fact that wild dogs wade or swim to traverse waterbodies, thus resulting in
slower, more tortuous movements. Besides this, our model also suggested that
dipsersers preferably realize shorter steps when moving through woodland, but
larger steps when moving across shrubs/grassland. This can likely be attributed
to resting periods during which dispersers seek shade and protection below the
woodland canopy. Since we did not include three way interactions, we could not
test this hypothesis.

\subsection{Simulation (80\%)}
% I need to better distinguish our approach from other IBMMs, maybe use the term
% iSSF-IBMM?
Based on the above outlined model results, we simulated 80'000 dispersers moving
2'000 steps across the landscapes of the KAZA-TFCA. Our dispersal simulation
thus comprised 160 Mio. simulated steps, which required five days of computation
on a modern desktop machine. This rather long simulation duration was primarily
owed to the massive extent considered (including the buffer ca. 1.8 Mio. km
\textsuperscript{2}) and the large number of dispersers simulated. Most
connectivity studies consider much smaller extents (e.g.
\citealp{Kanagaraj.2013, Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020})
and will therefore be able to attain faster simulation speeds. We also believe
that fewer simulated dispersers will often suffice, as the relative traversal
frequency of simulated individuals through randomly placed checkpoints in our
study area converged rather quickly (Appendices), yet this will vary depending
on the structure of the landscape and dispersal ability of the focal species.

A simulation-based approach as proposed in this article offers several striking
advantages over more traditional connectivity modeling techniques such as LCP
analysis or CT. In contrast to LCP analyses, for instance, an individual-based
simulation does not require to assume known endpoints, as endpoints emerge
naturally as the result of simulated dispersal trajectories. Liberating modelers
from the assumption of known endpoints is particularly important in dispersal
studies, as dispersers often venture into unfamiliar territory and are therefore
unlikely to know the destination of their journey. Additionally, without
pre-defined endpoints, movement corridors are no longer limited to certain
combinations of start- and endpoints, thereby permitting to detect movement
corridors that lead towards dead ends and insurmountable obstacles (i.e.
ecological traps).

IBMMs furthermore yield the advantage of an explicit representation of time.
This enables to answer questions such as: ``\textit{How long will it take a
disperser to move from A to B?}'' or \textit{Is it possible for a disperser to
move from A to B within X days?}, all undoubtedly interesting questions that
shift the focus from a structural to a more functional view on connectivity.
Similar questions cannot be answered using LCP methods or CT because the time
component is omitted from those models. An explicit representation of time in
IBMMs also yields opportunities for studying seasonal changes to connectivity
\cite{Zeller.2020}. With least-cost methods or circuit theory a dynamic analysis
of connectivity is impractical because the same analyses needs to be repeated
using different permeability surfaces (e.g. \citealp{Benz.2016, Osipova.2019}).
With IBMMs, on the other hand, simulated individuals can directly respond to a
dynamic environment within the model and the environment is allowed to change
``as the dispersers move''.

While an explicit representation of time yields several advantages, it also
requires that step lengths and turning angles are modeled properly
\citep{Kanagaraj.2013}, so that dispersal durations between habitat patches can
be reliably approximated. Accounting for and correctly rendering this movement
kernel is the main strength of ISSFs, which is why we believe that the framework
is particularly well suited for simulating animal dispersal and assessing
landscape connectivity. In principle, the ISSF framework allows to define an
infinite amount of movement kernels, each representing movement behavior under
different conditions \citep{Fieberg.2020}. It can therefore accommodate highly
diverse movement modes, including fast paced directional movements and slow
paced tortuous movements. Compared to models that disregard an animal's movement
kernel, this allows to model directional persistence, a movement characteristic
regularly observed in dispersing animals. To date, however, we only accounted
for 1. order autocorrelation. That is, we only considered correlations between
two subsequent steps. Modeling higher-order autocorrelation may be desirable,
yet will require more extensive data.

Overall, IBMMs enable a rich examination of the emergence of connectivity when
compared to the outputs of least-cost analysis or circuit theory. By generating
proper movement trajectories, a broad variety of summary statistics can be
derived. Specifically, by turning simulated trajectories into a network, network
theoretical metrics relevant to connectivity can be applied
\citep{BastilleRousseau.2018}. In contrast to least-cost path analysis and
circuit theory, a simulation-based approach permits studying inter-patch
connectivity. This facilitates the identification of asymmetrical links and
statements about source-sink dynamics \citep{Ferreras.2001, Revilla.2004,
Kanagaraj.2013}. Our results are also in concert with previous studies on
functional connectivity based on individual-based models that showed that
including the behavioral ecology of the target species and the landscape
structure are imperative when assessing connectivity \citep{Gustafson.1996,
Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004, Revilla.2008,
Kanagaraj.2013} rather than considering it purely a function of distance.

Besides the benefits of IBMMs, we also want to point out some caveats and not so
trivial modeling decisions. In particular, we will discuss five modeling
challenges: (1) number of simulated individuals, (2) location of source points,
(3) dispersal duration, (4) boundary behavior, and (5) individual variability.

(1) When simulating dispersal, the modeler needs to decide on the number of
simulated individuals. This decision involves the total (absolute) number of
simulated individuals across the study area, as well as the relative number of
simulated individuals per spatial entity (e.g. protected area, habitat patch,
source point). With regards to the total number of simulated individuals, a
larger number is usually desirable as it yields a more balanced view on
connectivity. However, this comes at the cost of computational efficiency,
implying that a trade-off needs to be managed. One simple solution to handle the
trade-off would be to simulate only as many individuals as are needed to achieve
convergence in the target metric. In our case, examination of the target metric
\textsf{traversal frequency} revealed that convergence was achieved with
relatively few individuals, such that additional dispersers contributed only
little additional information to our results. With regards to the relative
number of simulated individuals, we see two feasible approaches. First, one
could initiate dispersers in relation to population density at the respective
area. This would directly reflect the fact that population densities are not
homogeneous across space. (Say more) Alternatively, however, one could also
distribute dispersers homogeneously, but then weigh simulated trajectories by
the population density at the source patch (Say more). Because we lacked
reliable information on density estimates for the extent of the KAZA-TFCA, we
opted for a solution where dispersers were distributed homogeneously.

@@

(3) When employing IBMMs, one also needs to decide on meaningful dispersal
durations (number of simulated steps). We decided to simulate individuals for
2'000 steps, which is at the upper end of observed dispersal durations and may
have resulted in an overestimated representation of landscape connectivity.
Instead of enforcing homogeneous dispersal durations, one could also sample
dispersal durations from on observed dispersal events. Alternatively, after each
step, individuals could be faced with a probability to settle, where the
probability is contingent on the surrounding landscape and cumulative dispersal
distance traveled. Because we only observed few dispersal events and due to the
wild dog's ability to cover vast dispersal distance \citep{DaviesMostert.2012,
Masenga.2016, Cozzi.2020} we decided not to apply this approach. Regardless of
this, we think that in most cases it will be more insightful to simulate
relatively extensive dispersal events and only subsample afterwards.

@@

@@

% Downsides of connectivity?
Even though connectivity is generally thought to promote population viability,
it is also related to various aspects that may cause ecological damage, such as
increased human-wildlife conflicts or facilitated spread of deadly diseases.

% Conclusion
To this end, we have exemplified the use of ISSFs as a simple framework based on
which researchers can parametrize dispersers' habitat and movement preferences
and at the same time use the parametrized model to as an individual-based
movement model to simulate dispersal and examine landscape connectivity.
Furthermore, we suggest to analyse simulated trajectories using three
complementary maps, each focused on a different aspect of connectivity.
Ultimately, we hope to have sparked interest in the uprising framework of step
selection functions for investigating dispersal behavior and landscape
connectivity. Nevertheless, we do not attempt to dismiss the application of
traditional connectivity models such as least-cost methods or circuit theory by
any means. Rather, we propose to use simulations from ISSF-IBMMs as a simple but
powerful tool that complements earlier methods in an attempt provide a more
comprehensive understanding of landscape connectivity.

>>>>

% Previous Paper
\subsection{Previous Paper (90\%)}
In a previous study, we assessed landscape connectivity for dispersing African
wild dogs within the KAZA-TFCA using least-cost methods \citep{Hofmann.2021}.
Specifically, we fitted a basic habitat selection model and predicted a
permeability surface that we used to compute least-cost paths and corridors. We
now expand on this knowledge and use ISSFs to develop a more mechanistic
movement model of dispersing wild dogs (\Cref{GraphicalAbstract}). We employ the
model to simulate dispersers moving across the KAZA-TFCA and generate three
distinct connectivity maps, each shedding light onto a different aspect of
connectivity. With this work, we exemplify how ISSFs can be utilized for
dispersal simulations and we discuss several benefits of this approach over
traditional connectivity modeling techniques such as least-cost analysis and
circuit theory. Most importantly, simulations based on ISSFs provide a more
generic view on how connectivity emerges and to which degree connectivity
depends on the dispersal duration. In addition, by generating proper dispersal
trajectories, network theory can be applied to calculate network metrics that
are pertinent to connectivity analysis.

@@

Besides habitat covariates, we computed movement metrics that we used as
movement covariates in our models. Movement metrics were calculated for each
step and included the step length (\textsf{sl}), its natural logarithm
(\textsf{log(sl)}), and the cosine of the relative turning angle
(\textsf{cos(ta)}) (for details see \citep{Avgar.2016, Fieberg.2020}). Because
wild dogs follow a diurnal activity pattern \citep{Castello.2018}, we also coded
a binary variable (\textsf{LowActivity}) indicating whether a step was realized
during periods of low wild dog activity (17:00 to 09:00 local time) or high wild
dog activity (09:00 to 17:00 local time). Handling and manipulation of all data,
as well as all models and simulations were implemented with the statistical
software {\tt R}, version 3.6.6 \citep{R.2019}. Several helper functions were
written in {\tt C++} and imported into {\tt R} using the {\tt Rcpp} package
\citep{Eddelbuettel.2011, Eddelbuettel.2013}

@@

\subsection{Convergence (80\%)}
Our examination of the traversal frequency as a function of the number of
simulated dispersers shows that the mean traversal frequency stabilizes already
after very few simulations and changes only little when adding further
dispersers (\Cref{Convergence} (a) and (b)). While variability keeps decreasing
with additional dispersers, the marginal benefit of adding further dispersers
steeply decreases with a negative-exponential trend (\Cref{Convergence} (c)).

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{99_Convergence}
    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5)
    distributed randomly across the study area. The relative traversal frequency
    is plotted against the number of simulated individuals to visualize how
    quickly the metric converges to a steady state. (a) Replicated (100 times)
    relative traversal frequencies across three randomly chosen checkpoints as
    well as the corresponding 95\% prediction interval (PI). (b) Averaged
    relative traversal frequency across all checkpoints and replicates including
    a 95\% PI. (c) Width of the PI in relation to the number of simulated
    dispersers.}
    \label{Convergence}
  \end{center}
\end{figure}

@@

\section{Discussion}
\subsection{Short Summary (90\%)}
We used ISSFs to analyse data of dispersing wild dogs and to parametrize a fully
mechanistic movement model describing how dispersers move through the available
landscape. We employed the parametrized model as an individual-based movement
model to simulate 80'000 dispersing wild dogs moving 2'000 steps across the
extent of the KAZA-TFCA, the world's largest transboundary conservation area,
located in southern Africa. Based on simulated dispersal trajectories, we
derived three complementary maps, all geared towards a better understanding of
dispersal and landscape connectivity. These maps included a heatmap, revealing
frequently traversed areas, a betweenness-map, delineating critical dispersal
corridors, and a map of inter-patch connectivity, indicating the presence or
absence of functional links between national parks. We thereby showcase that
ISSFs offer a simple, yet powerful framework to parametrize movement models and
simulate dispersal to assess landscape connectivity. Importantly,
individual-based simulations from ISSFs overcome several conceptual shortcomings
inherent to more traditional connectivity modeling techniques, such as
least-cost path analysis and circuit theory.

\subsection{Movement Model (80 \%)}
Our movement model of dispersing wild dogs consisted of a habitat kernel, a
movement kernel, and their interactions. Together, the kernels in detail
described habitat and movement preferences of dispersing wild dogs, as well as
how movement behavior was affected by habitat conditions. Parameter estimates
from the habitat kernel revealed that dispersing wild dogs avoid water, prefer
its proximity, avoid woodland, prefer shrubs/grassland, and avoid areas
dominated by humans. These findings are consistent with findings from previous
studies on dispersing wild dogs \citep{DaviesMostert.2012, Masenga.2016,
Woodroffe.2019, Oneill.2020}, as well as with a base model that we developed
with a main focus on dispersers' habitat kernel \citep{Hofmann.2021}.

By expanding our base model by a proper movement kernel, we were able to model
several additional complexities inherent to dispersal. For instance, it is well
known that dispersers move with directional persistence and that their step
lengths are correlated with turning angles \citep{Morales.2004, Borger.2012}.
That is, larger steps usually imply smaller turning angles and vice versa. While
this behavior could be captured by sampling turning angles and step lengths from
copula probability distributions \citep{Hodel.2021}, the ISSF framework allowed
us to model similar behavior directly in the movement model. Besides accounting
for directional persistence and correlations between step lengths and turning
angles, we also allowed for interactions that rendered the fact that wild dogs
mainly move during the morning and evening hours, whereas they tend to rest
during the remainder of the day. The final model thus rendered a burst-like
movement behavior, where dispersers realize larger steps during the darker
morning and evening hours, but shorter steps during the rest of the day. While
we accounted for 1. order autocorrelation (i.e. correlation between two
subsequent steps), higher order autocorrelation is conceivable and might be
desirable to model. While ISSFs are conceptually capable of rendering such
behavior, it unfortunately requires vast amounts of GPS data that is not
intercepted by missing fixes, making it impractical to model higher-order
autocorrelation using ISSFs in reality.

By allowing interactions between habitat covariates and movement covariates, we
furthermore accounted for the fact that movement and habitat preferences are
interdependent. For example, the final model retained an interaction between
water cover and step length, showing that dispersers are more likely to realize
short steps (i.e. move slower) in areas covered by water. Likewise, the
parameter estimate for the interaction between water cover and turning angles
revealed that dispersers move less directional when moving across water bodies.
We believe that this behavior is owed to the fact that wild dogs wade or swim
when traversing waterbodies, thus resulting in slower, more tortuous movements.
Besides this, our model also suggested that dipsersers preferably realize
shorter steps when moving through woodland, but larger steps when moving across
shrubs/grassland. This can likely be linked to wild dogs' resting behavior, as
wild dogs usually use open areas to quickly move over long distances
\citep{Abrahms.2017} but seek shade and protection below the woodland canopy
when resting \citep{Creel.2002}. However, we did not consider three way
interactions and could therefore not test this hypothesis.

\subsection{Simulation (80\%)}
Based on the above described movement model, we simulated 80'000 dispersers
moving 2'000 steps across the landscapes of the KAZA-TFCA. This simulation of
160 Mio. steps required five days of computation on a modern desktop machine.
The long simulation time was primarily caused by the massive extent considered
(ca. 1.8 Mio. km \textsuperscript{2} when including the buffer) and the large
number of dispersers simulated. Most connectivity studies are limited to much
smaller extents (e.g. \citealp{Kanagaraj.2013, Clark.2015, McClure.2016,
Abrahms.2017, Zeller.2020}) and will therefore require shorter simulation times.
We also believe that fewer simulated dispersers will often suffice, as the
relative traversal frequency by simulated individuals through randomly placed
checkpoints converged rather quickly. In any case, the exact number of required
simulations will vary depending on the structure of the landscape and the
dispersal ability of the focal species.

A simulation-based approach as proposed in this article offers several
advantages over traditional connectivity modeling techniques such as LCP
analysis or CT. In contrast to LCP analysis, for instance, an individual-based
simulation does not require to assume known endpoints. Instead, each endpoint
emerges naturally from a simulated dispersal trajectory. This is particularly
valuable for dispersal studies, because dispersers often venture into unfamiliar
territory and are unlikely to know the destination of their journey
\citep{Elliot.2014, Abrahms.2017, Cozzi.2020}. A connectivity model that does
not require pre-defined endpoints also ensures that movement corridors are not
enforced between certain start- and endpoints and permits to reveal frequently
used routes that potentially lead into ecological traps \citep{Dwernychuk.1972,
VanDerMeer.2014}.

In comparison to LCP and CT approaches, IBMMs furthermore yield the advantage of
an explicit representation of time. The ability to render time enables to answer
questions such as: ``\textit{How long will it take a disperser to move from A to
B?}'' or \textit{Is it possible for a disperser to move from A to B within X
days?}. These are important questions that shift the focus from a structural to
a more functional view on connectivity. An explicit representation of time also
yields opportunities to study how connectivity depends on seasonal changes in
the environment (\textit{dynamic connectivity}; \citealp{Zeller.2020}). With LCP
or CT, a dynamic connectivity analysis is currently impractical because the same
analysis needs to be repeated for different permeability surfaces, each
resembling a snapshot in time for a different seasonal state of the environment
(e.g. \citealp{Benz.2016, Osipova.2019}). With IBMMs, on the other hand, the
environment is allowed to change dynamically ``as the dispersers move'' such
that simulated individuals can respond to seasonal factors directly within the
simulation.

While an explicit representation of time appears beneficial, it also requires
that step lengths and turning angles are modeled correctly
\citep{Kanagaraj.2013}, so that the time required to move between areas can be
reliably estimated. Correctly rendering movement behavior under varying
environmental conditions is one of the main strength of ISSFs \citep{Avgar.2016,
Prokopenko.2017, Fieberg.2020}, which is why we believe that the framework is
exceptionally well suited for simulating animal dispersal and assessing
landscape connectivity. In fact, the ISSF framework allows to define an infinite
amount of movement kernels, each representing movement preferences under
different environmental conditions \citep{Fieberg.2020}. ISSFs can therefore
accommodate highly diverse movement modes, including fast paced directional
movements and slow paced tortuous movements.

Whereas the results of LCP analysis and CT are usually restricted to single
``conductance'' maps, individual-based simulations can be processed using a
diverse suite of different approaches, each enabling to focus on a different
aspect of connectivity. By turning simulated trajectories into a network, for
instance, researchers can capitalize on concepts from network theory and
calculate network metrics that are pertinent to landscape connectivity
\citep{BastilleRousseau.2018}. Here, we have exemplified this by calculating the
betweenness metric, highlighting major movement corridors. Alternatively, can
learn about inter-patch connectivity and potential source-sink dynamics between
patches \citep{Ferreras.2001, Revilla.2004, Kanagaraj.2013} based on simulated
dispersal trajectories. Ultimately, by accounting for habitat- and movement
preferences of the focal species, simulations from ISSFs are capable of
rendering the behavioral ecology of the focal species in detail, thereby
permitting a more functional view on connectivity that is not solely driven by
the landscape structure and distances between patches \citep{Gustafson.1996,
Gardner.2004, Graf.2007, KramerSchadt.2004, Revilla.2004, Revilla.2008,
Kanagaraj.2013}.

\begin{figure}
  \includegraphics[width=\textwidth]{99_ModelingDecisions.png}
  \caption{Five major modeling decisions that a researcher needs to consider
  when simulating dispersers to assess landscape connectivity. (1) Number of
  simulated trajectories. (2) Handling individual variation: we used point
  estimates when simulating dispersers, assuming that there was no individual
  variation. Alternatively, however, one could draw preferences for each
  simulated individually based on model uncertainty. (3) Dispersal duration.
  While one could draw the number of simulated steps based on observed dispersal
  events, we opted for an alternative and simulated each individual for 2'000
  steps, which was at the upper end of observed dispersal durations. This allows
  to easily shorten the generated trajectories afterwards and to investigate the
  sensitivity of results with regards to the dispersal duration. (4) Boundary
  behavior. We allowed dispersers to leave and re-enter the main study area
  through a buffer zone with randomized covariate values. Alternatively, one
  could discard transgressing random steps, thereby forcing dispersers to remain
  within the study area or simply terminate the simulation, assuming the
  individual has disappeared. (5) Source point location. The location of source
  points should optimally be biologically informed. For our simulation, we
  placed source points within protected areas large enough to sustain viable
  wild dog populations. Conceivable alternatives include placing source points
  according to a habitat suitability model or based on abundance information.
  Importantly, one must also consider potential immigrants from outside the main
  study area.}
  \label{ModelingDecisions}
\end{figure}

Despite the benefits that ISSF simulations offer, we also want to confer some of
the non-trivial modeling decisions involved. In particular, we will discuss five
modeling decisions (\Cref{ModelingDecisions}): (1) number of simulated
individuals, (2) location of source points, (3) dispersal duration, (4) boundary
behavior, and (5) how to handle individual variability.

(1) When simulating dispersal, the modeler needs to decide on the number of
simulated individuals. This decision includes the \textit{absolute} number of
simulated individuals across the study area, as well as the \textit{relative}
number of simulated individuals per spatial entity (e.g. protected area, habitat
patch, source point). Because each additionally simulated individual adds
information about landscape connectivity and potential movement routes, a larger
absolute number of simulated individuals is always desirable, yet this comes at
the cost of computational efficiency. Consequently, a trade-off needs to be
managed. One solution to handle the trade-off would be to define a target metric
and to only simulate so many individuals until convergence in the target metric
is achieved. Here, we used the \textit{relative traversal frequency} across
checkpoints as target metric and revealed that convergence was achieved already
after few simulated individuals. With regards to the relative number of
simulated individuals, we see two feasible approaches. First, one could initiate
dispersers in relation to population density at the respective area. This would
directly reflect the fact that population densities are not homogeneous across
space. (Say more) Alternatively, however, one could also distribute dispersers
homogeneously, but then weigh simulated trajectories by the population density
at the source patch (Say more). Because we lacked reliable information on
density estimates for the extent of the KAZA-TFCA, we opted for a solution where
dispersers were distributed homogeneously.

@@

(3) When employing IBMMs, one also needs to decide on meaningful dispersal
durations (number of simulated steps). We decided to simulate individuals for
2'000 steps, which is at the upper end of observed dispersal durations and may
have resulted in an overestimated representation of landscape connectivity.
Instead of enforcing homogeneous dispersal durations, one could also sample
dispersal durations from observed dispersal events. Alternatively, after each
step, individuals could be faced with a probability to settle, where the
probability is contingent on the surrounding landscape and cumulative dispersal
distance traveled. Because we only observed few dispersal events and due to the
wild dog's ability to cover vast dispersal distance \citep{DaviesMostert.2012,
Masenga.2016, Cozzi.2020} we decided not to apply this approach. Regardless of
this, we think that in most cases it will be more insightful to simulate
relatively extensive dispersal events and only subsample afterwards.

@@

Approach also similar to \cite{Potts.2013}, yet they used regular SSFs and
examined steady sate UDs.

@@

% Conclusion
To this end, we have used data on dispersing wild dogs to exemplify how ISSFs
can be used to parametrize an individual-based movement model that is further
employed to simulate dispersal and examine landscape connectivity. We also
presented three complementary connectivity maps derived from simulated
trajectories, each focused on a different aspect of connectivity. Furthermore,
we discussed the potential advantages and disadvantages of the proposed
framework compared to traditional connectivity modeling techniques such as LCPs
and CT. With this article, we hope to have sparked interest in the uprising
framework of step selection functions for investigating dispersal behavior and
landscape connectivity. Nevertheless, we do not attempt to dismiss the
application of traditional connectivity models by any means. Rather, we propose
to use simulations from ISSF-models as a simple but powerful tool to provide a
more comprehensive understanding of dispersal and landscape connectivity.

%------------------------------------------------------------------------------
%  Further topics one could discuss
%------------------------------------------------------------------------------
% % Downsides of connectivity?
% Even though connectivity is in general thought to promote population viability,
% it is also related to various aspects that may cause ecological damage, such as
% increased connectivity into human-dominated areas or facilitated spread of
% deadly diseases \citep{Brearly.2013}. Nevertheless, information on the
% relationship between connectivity and the prevalence of deadly diseases in
% African wild dogs is currently lacking.
